---
title: "Seminar Work - Data Analysis"
author: "Antoine Thomas"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library import, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
library(xtable)
```

## Data Analysis

### Import of hotel prices lists

First, the scraped datasets with price information for selected hotels are imported. Then, the data records for all hotels are filtered so that they only contain price information for simple double bedrooms. In addition, care is taken to ensure that there is a maximum of one entry for each hotel each day. Thereby, the cheapest option is always chosen. This is primarily due to the fact that more expensive prices usually include various additional services, such as free cancellation up to the day of the start of the stay or meals such as breakfast. Thus, it can be assured that all prices cover a comparable range of services, regardless of the hotel. 
Subsequently, all data of the respective hotels are combined into one large data set.


```{r hotel price list import, echo=TRUE, message=FALSE, warning=FALSE}
# Data import
hotel_bb <- read_csv2(file = "B&B_Hotel_München-Hbf.csv")

# Filter for double bedrooms and keep the cheapest possible price per day
hotel_bb_clean <- hotel_bb %>%
  filter(room == "Doppelzimmer") %>%
  group_by(date, room) %>%
  slice(which.min(price))

# Further import process hidden........
```

```{r hidden hotel price list import, message=FALSE, warning=FALSE, include=FALSE}
# Data import
hotel_bold <- read_csv2(file = "Bold_Hotel_München.csv")
hotel_brunnenhof <- read_csv2(file = "Brunnenhof_City_Center.csv")
hotel_centro_mondial <- read_csv2(file = "Centro_Hotel_Mondial.csv")
hotel_demas <- read_csv2(file = "Hotel_Demas_München.csv")
hotel_gio <- read_csv2(file = "Hotel_GIO.csv")
hotel_mirabell <- read_csv2(file = "Hotel_Mirabell.csv")
hotel_city_center <- read_csv2(file = "Hotel_München_City_Center.csv")
hotel_munchen_city <- read_csv2(file = "Hotel_Munich_City.csv")
hotel_relexa <- read_csv2(file = "relexa_hotel_München.csv")

# Filter for double bedrooms and keep the cheapest possible price per day
hotel_bb_clean <- hotel_bb %>%
  filter(room == "Doppelzimmer") %>%
  group_by(date, room) %>%
  slice(which.min(price))

hotel_bold_clean <- hotel_bold %>%
  filter(room == "Bold Doppelzimmer") %>%
  group_by(date, room) %>%
  slice(which.min(price))

hotel_brunnenhof_clean <- hotel_brunnenhof %>%
  filter(room == "Doppelzimmer") %>%
  group_by(date, room) %>%
  slice(which.min(price))

hotel_centro_mondial_clean <- hotel_centro_mondial %>%
  filter(room == "Standard Doppelzimmer") %>%
  group_by(date, room) %>%
  slice(which.min(price))

hotel_demas_clean <- hotel_demas %>%
  filter(room == "Standard Doppelzimmer") %>%
  group_by(date, room) %>%
  slice(which.min(price))

hotel_gio_clean <- hotel_gio %>%
  filter(room == "Doppelzimmer") %>%
  group_by(date, room) %>%
  slice(which.min(price))

hotel_mirabell_clean <- hotel_mirabell %>%
  filter(room == "Superior Zimmer") %>%
  group_by(date, room) %>%
  slice(which.min(price))

hotel_city_center_clean <- hotel_city_center %>%
  filter(room == "Standard Zimmer") %>%
  group_by(date, room) %>%
  slice(which.min(price))

hotel_munchen_city_clean <- hotel_munchen_city %>%
  filter(room == "Doppel- oder Zweibettzimmer") %>%
  group_by(date, room) %>%
  slice(which.min(price))

hotel_relexa_clean <- hotel_relexa %>%
  filter(room == "Comfort Doppelzimmer") %>%
  group_by(date, room) %>%
  slice(which.min(price))

# Remove raw datas
remove(hotel_bb, 
       hotel_bold, 
       hotel_brunnenhof, 
       hotel_centro_mondial, 
       hotel_demas, 
       hotel_gio, 
       hotel_mirabell, 
       hotel_city_center, 
       hotel_munchen_city, 
       hotel_relexa)
```

```{r merging all datasets, echo=TRUE, message=FALSE, warning=FALSE}
# binding data of all hotels
hotel_data <- rbind(hotel_bb_clean,
        hotel_bold_clean,
        hotel_brunnenhof_clean,
        hotel_centro_mondial_clean,
        hotel_city_center_clean,
        hotel_demas_clean,
        hotel_gio_clean,
        hotel_mirabell_clean,
        hotel_munchen_city_clean,
        hotel_relexa_clean)
```

In the dataset `hotel_data` the high demand phase, the Oktoberfest, must now be noted. The 2023 event takes place from 16.09.2023 to 03.10.2023.
To do this, the function `checkOktoberfest` is created, which is given a date as parameter, and checks if this date is within period of the Oktoberfest.
After that a new attribute `oktoberfest` is created in `hotel_data`, which indicates with the help of the mentioned function, whether the Oktoberfest takes place on a specific day or not.

```{r add oktoberfest attribute, echo=TRUE, message=FALSE, warning=FALSE}
checkOktoberfest <- function(roomDate) {
  oktoberfestDateInterval = interval(ymd("2023-09-16"), ymd("2023-10-03"))
  roomDate %within% oktoberfestDateInterval
}

hotel_data <- hotel_data  %>%
  mutate(oktoberfest = unlist(map(.x = date, .f = checkOktoberfest)))
```

#### Price over time

In the following figures, the temporal price progression over the period from April 10, 2023 to December 31, 2023 is illustrated for the selected hotels. Vertical lines have been included on these to indicate the start and end of the Oktoberfest period. From these figures, it can be seen that all hotels retrieve the highest average prices of the year during the Oktoberfest period. Although there are clearly similar fluctuations in the pricing of hotel rooms at other times of the year, these cannot be compared with the extent of the Oktoberfest.

```{r price over time plots, echo=FALSE}
hotel_data %>%
  group_by(date) %>%
  summarise(price = mean(price)) %>%
  ggplot(aes(x = date, y = price)) +
  geom_line() +
  geom_vline(xintercept = date("2023-09-16"), colour = "red", alpha =.3) +
  geom_vline(xintercept = date("2023-10-03"), colour = "red", alpha = .3) +
  labs(x = "Date", y = "Room Rate", title = "Average price over time for a double bedroom in a hotel in the center of Munich") +
  theme(plot.title = element_text(face = "bold", size = "11"))
  
hotel_data %>%
  ggplot(aes(x = date, y = price)) +
  geom_line() +
  geom_vline(xintercept = date("2023-09-16"), colour = "red", alpha =.3) +
  geom_vline(xintercept = date("2023-10-03"), colour = "red", alpha = .3) +
  facet_wrap(~name) +
  theme_bw() +
  labs(x = "Date",y = "Room Rate", title = "Price over time for a double bedroom in 10 similar hotels in the center of Munich") +
  theme(plot.title = element_text(face = "bold",size = "11"))
```
