---
title: "Seminar Work - Data Analysis"
author: "Antoine Thomas"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Analysis

### Import of hotel prices lists

First, the scraped datasets with price information for selected hotels are imported. Then, the data records for all hotels are filtered so that they only contain price information for simple double bedrooms. In addition, care is taken to ensure that there is a maximum of one entry for each hotel each day. Thereby, the cheapest option is always chosen. This is primarily due to the fact that more expensive prices usually include various additional services, such as free cancellation up to the day of the start of the stay or meals such as breakfast. Thus, it can be assured that all prices cover a comparable range of services, regardless of the hotel. 
Subsequently, all data of the respective hotels are combined into one large data set.


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
